---
title: "Homework 10"
author: "Moiya Josephs"
date: "4/7/2022"
output: 
  html_document:
    toc: true
    toc_float: true
    toc_depth: 4
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidytext)
library(textdata)
library(janeaustenr)
library(dplyr)
library(stringr)
library(syuzhet)
library(ggplot2)
library(tidyr)
library(rvest)
library(knitr)
```

# Overview

In this assignment, I used the primary example code from chapter 2 working in this R Markdown document. In the chapter the authors used three sentiment lexicons, *Afinn*, *bing*, and *nrc*.
After we were to work with a different corpus of our choosing, and incorporate at least one additional sentiment lexicon. I chose the new Supreme Court Justice Ketanji Brown Jackson’s opening speech and found the sentiment based on the loughron lexicon.

# Textbook example
This section goes over a brief part of the example given in the textbook. 
## Sentiment Lexicons

### Afinn

```{r}
#Source: "Text Mining with R: A Tidy Approach" Chapter 2
knitr::kable(head(get_sentiments("afinn")))
```
### Bing

Bing categorizes words into positive and negative categories.

```{r}
#Source: "Text Mining with R: A Tidy Approach" Chapter 2
knitr::kable(head(get_sentiments("bing")))
```
### NRC  

The NRC lexicon categorizes words that are binary into categories that are positive, negative, anger, anticipation, disgust, fear, etc.

```{r}
#Source: "Text Mining with R: A Tidy Approach" Chapter 2

knitr::kable(head(get_sentiments("nrc")))

```

In the textbook example, the code found the sentiment on Jane austen boosk. The data when tidy can be inner joined with the specific sentiment lexicons.
```{r}
#Source: "Text Mining with R: A Tidy Approach" Chapter 2

tidy_books <- austen_books() %>% group_by(book) %>% mutate(linenumber = row_number(), chapter = cumsum(str_detect(text, regex("^chapter [\\divxlc]", ignore_case = TRUE)))) %>% ungroup() %>% unnest_tokens(word,text)

knitr::kable(head(tidy_books))
```

Next the textbook used the NRC lexicon to perform sentiment analysis and filter to find the joy words. Below the tidy_books represents the where in the book Emma there is words NRC defines as joy.
```{r}
#Source: "Text Mining with R: A Tidy Approach" Chapter 2

nrc_joy <- get_sentiments("nrc") %>% filter(sentiment == "joy")
tidy_books %>% filter(book == "Emma") %>% inner_join(nrc_joy) %>% count(word, sort=TRUE)

knitr::kable(head(nrc_joy))

```

```{r}
#Source: "Text Mining with R: A Tidy Approach" Chapter 2


jane_austen_sentiment <- tidy_books %>% inner_join(get_sentiments("bing")) %>% count(book, index = linenumber %/% 80, sentiment) %>% pivot_wider(names_from = sentiment, values_from = n, values_fill = 0) %>% mutate(sentiment = positive - negative)

ggplot(jane_austen_sentiment, aes(index, sentiment, fill = book)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~book, ncol = 2, scales = "free_x")
```
## Comparing the three sentiment dictionaries

```{r}
#Source: "Text Mining with R: A Tidy Approach" Chapter 2
pride_prejudice <- tidy_books %>% 
  filter(book == "Pride & Prejudice")

knitr::kable(head(pride_prejudice))

```
```{r}
#Source: "Text Mining with R: A Tidy Approach" Chapter 2

afinn <- pride_prejudice %>% 
  inner_join(get_sentiments("afinn")) %>% 
  group_by(index = linenumber %/% 80) %>% 
  summarise(sentiment = sum(value)) %>% 
  mutate(method = "AFINN")

bing_and_nrc <- bind_rows(
  pride_prejudice %>% 
    inner_join(get_sentiments("bing")) %>%
    mutate(method = "Bing et al."),
  pride_prejudice %>% 
    inner_join(get_sentiments("nrc") %>% 
                 filter(sentiment %in% c("positive", 
                                         "negative"))
    ) %>%
    mutate(method = "NRC")) %>%
  count(method, index = linenumber %/% 80, sentiment) %>%
  pivot_wider(names_from = sentiment,
              values_from = n,
              values_fill = 0) %>% 
  mutate(sentiment = positive - negative)

bind_rows(afinn, 
          bing_and_nrc) %>%
  ggplot(aes(index, sentiment, fill = method)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~method, ncol = 1, scales = "free_y")
```


```{r}
#Source: "Text Mining with R: A Tidy Approach" Chapter 2
knitr::kable(get_sentiments("bing") %>% 
  count(sentiment))
```

## Most common positive and negative words
```{r}
# Source: "Text Mining with R: A Tidy Approach" Chapter 2
bing_word_counts <- tidy_books %>%
  inner_join(get_sentiments("bing")) %>%
  count(word, sentiment, sort = TRUE) %>%
  ungroup()

knitr::kable(head(bing_word_counts))
```
```{r}
# Source: "Text Mining with R: A Tidy Approach" Chapter 2
bing_word_counts %>%
  group_by(sentiment) %>%
  slice_max(n, n = 10) %>% 
  ungroup() %>%
  mutate(word = reorder(word, n)) %>%
  ggplot(aes(n, word, fill = sentiment)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~sentiment, scales = "free_y") +
  labs(x = "Contribution to sentiment",
       y = NULL)
```
<hr>

# Sentiment on a new Corpus


## Scrape website 

To find the speech I scraped the PBS website of the transript of Judge Jackson's opening remarks. Then I saved it to value `speech`.

```{r}
speech_website <- read_html("https://www.pbs.org/newshour/politics/read-the-full-text-of-supreme-court-nominee-ketanji-brown-jacksons-opening-remarks")
speech <- speech_website %>%
html_nodes("p") %>%
html_text()
```



## Syuzhet

Syuzhet is a package with its own lexicon. In the package it defines that its sentiments function returns a data frame where each row represents a sentence from the original file. The different columns are: "anger", "anticipation", "disgust", "fear", "joy", "sadness", "surprise", "trust", "negative", "positive." 

Instead of `get_sentiments` the function is `get_sentiment`. For each sentence it gives a numeric value of the sentiment. Shown in the output below. 

```{r}
get_sentiment(speech[3:24])
```

With `get_nrc_sentiment` it calculates eight different emotions and the sentiments in the text file. 

```{r}

knitr::kable(get_nrc_sentiment(speech[3:24]))
```


```{r}
s_v <- get_sentences(speech[3:24])
s_v_sentiment <- get_sentiment(s_v)
s_v_sentiment
```


## Loughran 

English sentiment lexicon created for use with financial documents. This lexicon labels words with six possible sentiments important in financial contexts: "negative", "positive", "litigious", "uncertainty", "constraining", or "superfluous".


```{r}
knitr::kable(get_sentiments("loughran") %>% count(sentiment, sort = TRUE) )
```

The speech is mainly in the body from 3 to 24, so I slice the body of the webpage and save it to `tidy_speech`. Now I need to extract all the words, I strip the empty space and convert it into a data frame. Each row is a single word in the speech.

```{r}
tidy_speech <- speech[3:24]

tidy_speech_words <- unlist(as.list(strsplit(tidy_speech, " ")))
rowNumber <- seq(1:length(tidy_speech_words))
words.df <- data.frame(rowNumber, tidy_speech_words)
names(words.df) <- c("rowNumber","word")
```

Just like the textbook example, I inner join the words in the loughran sentiment to the speech.
```{r}
speech_sentiment_loughran <- words.df %>% inner_join(get_sentiments("loughran"))
```


To count the amount of words per sentiment type I again followed the textbook. Inner join the words from the speech with the lexicon, then count the words by the sentiment and sort it. Next I graphed it using ggplot. 
```{r}
lang_word_counts <- words.df %>%
  inner_join(get_sentiments("loughran")) %>%
  count(word, sentiment, sort = TRUE) %>%
  ungroup()
lang_word_counts %>%
  group_by(sentiment) %>%
  slice_max(n, n = 10) %>% 
  ungroup() %>%
  mutate(word = reorder(word, n)) %>%
  ggplot(aes(n, word, fill = sentiment)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~sentiment, scales = "free_y") +
  labs(x = "Contribution to sentiment",
       y = NULL)
```
Here you can see uncertainty was the largest sentiment type in the speech followed by positive. Loughran was able to identify some litigious words, which is expected from a speech for a potential Supreme Court judge. 


# Conclusion  

When using lexicons it is important to note how binary the scoring of the words are and where the source of the text came from. Loughran for example was based on and mainly used for financial documents. The textbook shows how different lexicons can distribute sentiments differently. A possible extention to the new corpus is to compare the sentiment with different lexicons like the textbook did. Context also matters as it can change the meaning of words than just looking at them atomically. 


# References

Silge, Julia, and David Robinson. "Text Mining with R: A Tidy Approach". O'Reilly Media, 2017

